TBD

### 2020-01-13 manuscript/manuscript.Rmd
first draft of incorporating reviewer comments

*Keywords*: adaptive landscape, ecological networks, eco-evolutionary dynamics, natural selection, host-parasitoid, <!--complexity, biodiversity,-->community context, extinction.

@@

The loss of biodiversity<!--Global change--> is altering the structure of ecological networks; however, we are currently in a poor position to predict how these altered communities will affect the evolutionary potential of remaining populations. Theory on adaptive landscapes provides a framework for predicting how selection constrains phenotypic evolution, but often treats the network context of evolving populations as a ``black box''. Here, we integrate ecological networks and adaptive landscapes to examine how changes in food-web structure shape evolutionary constraints. We conducted a field experiment that removed<!--simulated the extinction of--> a guild of larval parasitoids that imposed direct and indirect selection pressures on an insect herbivore. We then measured herbivore survival as a function of three key phenotypic traits to estimate directional, quadratic, and correlational selection gradients in each treatment. We used these selection gradients to characterize the slope and curvature of the adaptive landscape to understand the direct and indirect effects of consumer loss on evolutionary constraints. We found that the number of traits under directional selection increased with the removal of larval parasitoids, indicating greater selective constraints on the trajectory of evolutionary change. Similarly, we found that the removal of larval parasitoids altered the curvature of the adaptive landscape in such a way that tended to decrease the evolvability of the traits we measured in the next generation.<!--In contrast, the adaptive landscape was more neutral in the original food web because different parasitoid guilds impose different selection pressures, minimizing relative fitness differences among phenotypes.--> Our results suggest that the loss of trophic interactions <!--due to consumer extinctions -->can impose greater constraints on phenotypic evolution. This indicates that the simplification of ecological communities may constrain the adaptive potential of remaining populations to future environmental change.

@@

The adaptive landscape provides a powerful framework for understanding how natural selection has shaped the evolution of biodiversity ---from genes to phenotypes to species [@Wright1931; @Simpson1944; @Arnold2001]. More than a metaphor, the adaptive landscape links quantitative genetic and phenotypic variation to evolution by natural selection [@Lande1979; @Arnold1984applications; @Arnold1984theory], giving insight to the direct and indirect effects of selective constraints [@Arnold1992]. Ecological interactions often play a key role in shaping selective constraints, as evidenced by the role of antagonistic and mutualistic interactions in driving evolutionary change [@Schluter2000; @Abrams2000; @Bronstein2006]. Although there is clear evidence that species interactions can shape the adaptive landscape, we still have a poor understanding of how the adaptive landscape is shaped by a community of interacting species [@McPeek2017; @terHorst2018]. Resolution on how change in ecological communities shape phenotypic evolution is urgently needed though, given the rapid losses of biodiversity we are observing in the Anthropocene [@Scheffers2016].

Ecological networks, such as food webs describing who-eats-whom, provide an explicit representation of the direct and indirect effects that emerge in a community of interacting species [@Bascompte2014; @McCann2012]. Here, we integrate ecological networks and adaptive landscapes to understand how ecological communities constrain evolutionary change. The direct and indirect effects of natural selection can be inferred by quantifying the slope and curvature of the adaptive landscape [@Arnold1992]. For example, the slope is determined by directional selection gradients acting on each phenotypic trait and influences the trajectory of evolutionary change [@Lande1979;@Arnold1992]. Selective constraints on evolution increase with the number of traits under directional selection, as this diminishes the number of optimal solutions [@Arnold2003]. The curvature of the adaptive landscape is governed by nonlinear selection gradients and acts to indirectly constrain evolution through its effect on genetic constraints [@Arnold1992; @Hansen2008]. Genetic constraints are largely governed by a population's G-matrix ---the additive genetic variances and covariances between traits [@Hansen2008]. Stabilizing selection acts to erode genetic variance in traits, which can impose a constraint on the ability of those traits to respond to future selection [@Hansen2008].<!-- In general, genetic constraints will increase with the number of traits under directional or stabilizing selection, as this will decrease the additive genetic variance in those traits [@Hansen2008].--> Correlational selection may alter the genetic covariance between traits, which may facilitate or hinder future adaptation depending on the pattern of selection on those traits [@Hansen2008].<!--Genetic constraints may also increase with the number of trait combinations under correlational selection, as this type of selection decreases the evolutionary independence of traits [@Hansen2008].--> If we want to predict how ecological communities shape phenotypic evolution, we must understand how ecological networks shape the adaptive landscape.

The loss of biodiversity is altering the structure of ecological networks<!--communities-->, which may influence evolutionary constraints in a number of ways. For example, in a <!--multitrophic community or -->food web, if different consumers impose directional selection on different traits of a shared resource, then a more diverse consumer community may constrain evolution by increasing the number of traits under selection. Alternatively, if consumers impose selection on different values of a trait, then their selective effects would cancel each other out in more diverse communities. To examine these different possibilities, we conducted a field experiment that removed a consumer guild that directly and indirectly impacts fitness of an abundant insect herbivore (*Iteomyia salicisverruca*, Family Cecidomyiidae; fig. \ref{fig:Conceptual}). The larvae of this herbivore induce tooth-shaped galls when they feed on the developing leaves of willow trees [*Salix* sp., @Russo2006]. These galls protect larva from attack by generalist predators (e.g. ants, spiders), but they suffer high mortality from egg and larval parasitoids [@Barbour2016]. We manipulated food-web structure by either removing larval parasitoids (removal food web) or allowing both egg and larval parasitoids to impose selection on gall midge traits (original food web, fig. \ref{fig:Conceptual}). The relative abundance of these larval parasitoids is much lower than the egg parasitoid [@Barbour2016], and thus represents a plausible extinction scenario for this system. <!--While this manipulation simulates an extreme co-extinction event, the likelihood of such co-extinction events is increasing in the Anthropocene. -->Larval parasitoids also impose indirect effects on gall midge fitness through intraguild predation on the egg parasitoid (fig. \ref{fig:Conceptual}). We applied modern statistical methods to quantify how changes in food-web structure altered the slope and curvature of the gall midge's adaptive landscape. Taken together, our study gives insight to how the loss of biodiversity may constrain the evolution of interacting populations.

@@

We setup our food-web manipulation across 128 plants soon after galls began developing on willows in early June of 2013. These 128 plants came from eight different plant genotypes that spanned the range of trait variation observed in this willow population [@Barbour2015]. For the original food web (eight replicates per genotype), we used flagging tape to mark 14 galled leaves per plant (~30 larvae), allowing the full suite of egg and larval parasitoids to impose selection. Marking galls with flagging tape ensured that all galls had similar phenology when we collected galls later in the season. For the removal food web, we enclosed 14 galled leaves with 10x15cm organza bags (ULINE, Pleasant Prairie, WI, USA) to exclude three parasitoid species that attack during larval development. This treatment did not exclude the egg parasitoid *Platygaster* sp., which attacks prior to gall initiation [larva initiate gall development in Cecidomyiid midges: @Gagne1989]. It was not possible to reciprocally manipulate parasitoid attack (i.e. exclude egg parasitoid, but allow larval parasitoids), because it was not possible to identify midge oviposition sites prior to gall formation. In late August, we collected marked and bagged galls from each plant, placed them into 30 mL vials and kept them in the lab for 4 months at room temperature. We then opened galls under a dissecting scope and determined whether larvae survived to pupation (our measure of fitness) or were parasitized. We did not include other sources of mortality in our analyses, such as early larval death, as they could influence the expression of the gall phenotype and confound estimates of selection. <!--Since we were interested in selection imposed by parasitoids, we excluded other sources of mortality (`r round((mean(full_df$early.larval.death) + mean(full_df$unknown_mortality)), 2)*100`% mortality from other sources). Another reason why we excluded other sources of mortality, such as early larval death, is that they can influence the expression of the gall phenotype and confound estimates of selection.--> For the food-web treatment that excluded larval parasitoids, we further restricted our data by removing any incidental instances of parasitism by a larval parasitoid. This represented less than 3\% of the observations in this food-web treatment and allowed us to focus our inferences of selection on those imposed by the egg parasitoid. Our final dataset contains survival estimates for `r no_larvae` larvae from `r no_galls` galls and `r no_plants` plants.

@@

We collected data on three different traits that we expected to influence larval survival based on previous work in this system [@Barbour2016] and other work with gall midges [@Weis1983; @Heath2018]. First, we measured gall diameter as the size of each gall chamber to the nearest 0.01 mm at its maximum diameter (perpendicular to the direction of plant tissue growth). Previous work in this system has shown that larger galls are associated with higher survival [@Barbour2016]. Second, we measured clutch size by counting the number of chambers in each gall [@Weis1983; @Heath2018]. All larvae collected from the same multi-chambered gall were scored with the same clutch size. Third, we measured oviposition preference as the density of larvae observed on a plant in an independent survey. We did this by randomly sampling five branches per tree and counting the number of individual gall chambers (number of larvae). We then converted these counts to a measure of larval density per 100 shoots by counting the number of shoots on the last branch we sampled. All larvae collected from the same plant were scored with the same oviposition preference. Measuring larval densities on plants in the field is a common method for measuring oviposition preference [@Gripenberg2010]; however, caution must be taken in inferring 'preference' as larval densities can be influenced by processes other than preference [@Singer1986]. Fortunately, a couple features of our study system suggest that larval densities may be a good proxy for oviposition preference. For example, since our data comes from a randomized placement of plant genotypes in a common garden, there is no consistent bias in which plant genotypes females are exposed to while searching for oviposition sites. Also, egg predation is a minor source of mortality for galling insects in general [@Hawkins1997]; therefore, we do not expect any prior egg predation to bias our estimates of observed larval densities.

## Quantifying the Adaptive Landscape

Inferring adaptive landscapes assumes that trait distributions are multivariate normal [@Lande1983]. 
To approximate this assumption, we log-transformed clutch size and square-root transformed oviposition preference. We then scaled all phenotypic traits (mean=0 and SD=1) across treatments prior to our analyses (detailed below) to ensure that our estimates of selection were comparable across traits and with other studies.

Our analysis consisted of four parts. First, we used generalized linear mixed models (GLMM) to quantify selection surfaces \textemdash linear and nonlinear relationships between absolute fitness ($W$) and standardized phenotypic traits ($i$) of individuals \textemdash in each food-web treatment. Second, we translated selection surfaces into the scale of relative fitness ($w$) in order to calculate standardized selection gradients. Third, we used our estimates of directional selection gradients to characterize the slope of the adaptive landscape, which we used to measure selective constraints. Finally, we estimated the curvature of the adaptive landscape and used a simulation to explore its effects on the adaptive potential of the gall midge population in the next generation.

**Selection surfaces**: Since larval survival was our measure of absolute fitness, we used a GLMM that assumed a binomial error distribution (and logit-link function). To approximate the selection surface, we modelled larval survival as a function of food-web treatment as well as linear ($\alpha_i$), quadratic ($\alpha_{ii}$), and linear interactions ($\alpha_{ij}$) between each trait. We also allowed these selection surfaces ($\alpha$) to vary between food-web treatments. Note that to obtain valid estimates of linear selection surfaces, we removed nonlinear terms prior to estimating linear relationships [@Lande1983]. Other approaches have been advocated for approximating selection surfaces [@Schluter1988]; however, our approach enables us to calculate selection gradients, and thus is more appropriate for approximating the adaptive landscape [@Arnold2003]. To account for the nonindependence of clutch size (measured at gall level) and oviposition preference (measured at plant level) as well as any independent effects of willow genotype on larval survival, we modelled gall ID nested within plant ID nested within genotype ID as random effects. Although statistical models with random effects are not common in analyses of natural selection, we think that modelling random effects can mitigate biased estimates of selection due to environmental covariances between traits and fitness [@Rausher1992]. Since our end goal was to characterize the relationship between mean trait values and mean fitness (adaptive landscape), we assumed the mean value of our random effects (i.e. setting them to zero) when calculating selection surfaces. We then used parametric bootstrapping (1,000 replicates) to estimate the effect of food-web treatment on larval survival as well as selection surfaces in each food-web treatment. To determine whether trait-fitness relationships differed between food-web treatments, we calculated the difference in bootstrapped replicates between treatments.

**Selection gradients**: Selection gradients cannot be estimated directly from GLMMs of selection surfaces because the response is in terms of absolute fitness and the coefficients are on a nonlinear scale. For example, the coefficients in the above model measure the change in the logarithm of the odds of surviving (i.e. $\text{ln}\{W(z)/[1-W(z)]\}$) associated with 1 SD change in a trait with all other traits held fixed at their mean. Therefore, we used the method developed by @Janzen1998 to translate selection surfaces from the above model into the scale of relative fitness in order to calculate directional ($\beta_i$), quadratic ($\gamma_{ii}$), and correlational ($\gamma_{ij}$) selection gradients. Briefly, this method calculates the average gradient of selection surfaces by multiplying the average of $W(z)[1-W(z)]$ by each regression coefficient (e.g. $\alpha_i$, $\alpha_{ii}$, or $\alpha_{ij}$). We then divided this average gradient by the mean fitness ($\bar W$) to put it on the scale of relative fitness, and thus interpretable as a selection gradient. We estimated selection gradients separately for each food-web treatment. We also determined whether selection gradients differed between food-web treatments by calculating the difference in bootstrapped replicates between treatments. Note that we doubled all quadratic terms prior to calculating selection gradients to put them on the same scale as estimates of directional and correlational selection [@Stinchcombe2008].

**Selective constraints**: <!--We measured evolutionary constraints by inspecting the slope and curvature of the adaptive landscape. -->The number of selective constraints is determined by the slope of the adaptive landscape, which in our study corresponds to:

$$\textbf{Slope} = \begin{pmatrix} \beta_{\text{Diam}} \\ \beta_{\text{Clutch}} \\ \beta_{\text{Pref}} \end{pmatrix} $$
where each $\beta_i$ corresponds to the directional selection gradient acting on each trait. By comparing the number of directional selection gradients that show clear evidence of contributing to the slope (i.e. 95%  CI does not overlap zero) in our food-web treatments, we can infer the effect of food-web structure on selective constraints. This gives insight as to whether there is a specific combination of phenotypes favored by natural selection (i.e. more traits under directional selection), or whether multiple trait combinations have comparable fitness (i.e. less traits under directional selection).

**Genetic constraints**: The indirect effects of selection on the G-matrix ($\Delta\text{G}=\text{G}(\gamma - \beta \beta^\text{T})\text{G}$) are governed by the curvature of the adaptive landscape ($\text{C}=\gamma - \beta \beta^\text{T}$), which in our study corresponds to:

@@

where each $\gamma_{ii}$ (diagonal) corresponds to the quadratic selection gradient acting on a trait, and each $\gamma_{ij}$ (off-diagonal) corresponds to the correlational selection gradient acting on a particular trait combination. Note that we ommitted the upper triangle of each matrix for clarity since it is simply the reflection of the lower triangle. Subtracting these two matrices results in the curvature matrix of the adaptive landscape:

@@

The G-matrix itself is a complex structure, but has a clear theoretical link to the evolvability of phenotypic traits [@Hansen2008]. <!--We used evolvability as our measure of genetic constraint, as it has a solid foundation in theory of quantitative genetics and adaptive landscapes [@Hansen2008]. -->Evolvability measures the ability for a trait to evolve toward a given direction of selection [@Hansen2008]. In the absence of knowledge about the direction of selection that a population will actually experience in the next generation, we can compute its average evolvability over random selection gradients [@Hansen2008; @Melo2015]. By computing the average evolvability (here, we used `r n_random_gradients` random $\beta$s) for each of `r n_Gmats` random G-matrices, we can then assess how changes in the curvature matrix alter the evolutionary potential of the associated traits. We report the distribution of these effect sizes, rather than conduct a statistical test, because the number of replicates in a simulation can be arbitrarily high, thus making any effect size statistically significant [@White2014].

@@

Once parasitized, the gall phenotype becomes the phenotype of the egg parasitoid. This phenotype may influence the egg parasitoid's survival in the face of larval parasitoids, and thus experiences selection. Our food-web manipulation allows us to measure selection imposed by larval parasitoids on the phenotype of egg parasitoids. Using the same models as described above, we substituted egg parasitism as our response variable to quantify selection surfaces and selection gradients acting on the egg parasitoid. Note that we cannot test the effect of food-web structure on the egg parasitoid's adaptive landscape ---we can only estimate selection imposed by larval parasitoids. This comparison is still useful though in determining the extent to which the loss of consumers may have indirect evolutionary effects by altering selection on multiple interacting populations.

>>>>

*Keywords*: **fitness landscape**, adaptive landscape, ecological networks, eco-evolutionary dynamics, natural selection, host-parasitoid, <!--complexity, biodiversity,-->community context, extinction.

@@

The loss of biodiversity<!--Global change--> is altering the structure of ecological networks; however, we are currently in a poor position to predict how these altered communities will affect the evolutionary potential of remaining populations. Theory on **fitness** landscapes provides a framework for predicting how selection constrains phenotypic evolution, but often treats the network context of evolving populations as a ``black box''. Here, we integrate ecological networks and **fitness** landscapes to examine how changes in food-web structure shape evolutionary constraints. We conducted a field experiment that removed<!--simulated the extinction of--> a guild of larval parasitoids that imposed direct and indirect selection pressures on an insect herbivore. We then measured herbivore survival as a function of three key phenotypic traits to estimate directional, quadratic, and correlational selection gradients in each treatment. We used these selection gradients to characterize the slope and curvature of the **fitness** landscape to understand the direct and indirect effects of consumer loss on evolutionary constraints. We found that the number of traits under directional selection increased with the removal of larval parasitoids, indicating greater selective constraints on the trajectory of evolutionary change. Similarly, we found that the removal of larval parasitoids altered the curvature of the **fitness** landscape in such a way that tended to decrease the evolvability of the traits we measured in the next generation.<!--In contrast, the adaptive landscape was more neutral in the original food web because different parasitoid guilds impose different selection pressures, minimizing relative fitness differences among phenotypes.--> Our results suggest that the loss of trophic interactions <!--due to consumer extinctions -->can impose greater constraints on phenotypic evolution. This indicates that the simplification of ecological communities may constrain the adaptive potential of remaining populations to future environmental change.

@@

The **fitness** landscape provides a powerful framework for understanding how natural selection has shaped the evolution of biodiversity ---from genes to phenotypes to species [@Wright1931; @Simpson1944; @Arnold2001]. More than a metaphor, the **fitness** landscape links quantitative genetic and phenotypic variation **in multiple traits** to evolution by natural selection [@Lande1979; @Arnold1984applications; @Arnold1984theory]<!--, giving insight to the direct and indirect effects of selective constraints [@Arnold1992]-->. Ecological interactions often play a key role in shaping natural selection<!--selective constraints-->, as evidenced by the role of antagonistic and mutualistic interactions in driving evolutionary change [@Schluter2000; @Abrams2000; @Bronstein2006]. Although there is clear evidence that species interactions can shape the **fitness** landscape, we still have a poor understanding of how the **fitness** landscape is shaped by a community of interacting species [@McPeek2017; @terHorst2018; **@Hui2018**]. Resolution on how change in ecological communities shape phenotypic evolution is urgently needed though, given the rapid losses of biodiversity we are observing in the Anthropocene [@Scheffers2016].

Ecological networks, such as food webs describing who-eats-whom, provide an explicit representation of the direct and indirect effects that emerge in a community of interacting species [@Bascompte2014; @McCann2012]. Here, we integrate ecological networks and **fitness** landscapes to understand how ecological communities constrain evolutionary change **[@Hui2018]**. The <!--direct and indirect -->effects of natural selection can be inferred by quantifying the slope and curvature of the **fitness** landscape **for multiple phenotypic traits** [@Arnold1992]. <!--For example, t-->The slope is determined by the **strength of** directional selection <!--gradients -->acting on each trait**,which** influences the trajectory of evolutionary change [@Lande1979;@Arnold1992]. **This fact is made clear by the 'Lande equation', $\Delta \bar z = \beta\text{G}$, where $\Delta \bar z$ is a column vector of the average change in each trait between generations, $\beta$ is a column vector of directional selection gradients acting on each trait (i.e. slope), and $\text{G}$ is the additive genetic (co)variance matrix for these traits (i.e. G-matrix). @Grant1995 used their knowledge of the slope of the fitness landscape for multiple traits related to body and beak size in Darwin's finches <!--, in addition to the G-matrix of those traits, -->to accurately predict the trajectory of evolution in these traits in response to different episodes of selection imposed by drought.**
<!--For example, a drought in 1976-1977 on Daphne Major in the Gal\'apagos imposed directional selection for large weight, wing length, and bill depth in medium ground finches, and this equation predicted well the evolutionary response in all three of these traits in the next generation (Grant and Grant 1995).
This equation has been used to predict short-term evolutionary change in Darwin's finches (Grant and Grant 1995).-->
<!--Selective constraints on evolution increase with the number of traits under directional selection, as this diminishes the number of optimal solutions [@Arnold2003].-->
The curvature of the **fitness** landscape is governed by the **strength of stabilizing, disruptive, and correlational selection acting on each trait<!--directional, quadratic, and correlational <!--nonlinear selection gradients-->, which <!--and acts to to indirectly constrain evolution -->can alter the potential for evolutionary change through its effect on <!--genetic constraints [@Arnold1992; @Hansen2008]. Genetic constraints are largely governed by--><!-- a population's -->the G-matrix** <!-----the additive genetic variances and covariances between traits -->[@Hansen2008]. 
For example, stabilizing selection acts to erode genetic variance in traits, which can impose a constraint on the ability of those traits to respond to future selection [@Hansen2008]. **In contrast, disruptive selection for different trait values acts to increase genetic variance in traits, thus increasing the capacity for future adaptation [@Bolnick2008]**.<!-- In general, genetic constraints will increase with the number of traits under directional or stabilizing selection, as this will decrease the additive genetic variance in those traits [@Hansen2008].--> Correlational selection alters the genetic covariance between traits, which may facilitate or hinder future adaptation depending on the pattern of selection on those traits **and the structure of the G-matrix** [@Brodie1992].<!--Genetic constraints may also increase with the number of trait combinations under correlational selection, as this type of selection decreases the evolutionary independence of traits [@Hansen2008].--> If we want to predict how ecological communities shape phenotypic evolution, we must understand how ecological networks shape the **fitness** landscape **of interacting populations**.

The loss of biodiversity is altering the structure of ecological networks<!--communities-->**[@Landi2018ComplexityStability]**, which may influence **the slope and curvature of the fitness landscape**<!-- evolutionary constraints -->in a number of ways. **Consider first the slope of the fitness landscape <!--For example, -->in a <!--multitrophic community or -->food web. If different consumers impose directional selection on different traits of a shared resource, then a more diverse consumer community <!--may constrain evolution by -->**would** increase the number of traits under selection**, which may constrain the trajectory of evolution toward a specific phenotype**. Alternatively, if consumers impose selection on different values of a trait, then their selective effects would cancel each other out in more diverse communities**, which would allow for a greater diversity of phenotypes. Now consider the curvature of the fitness landscape. If consumers impose selection on different ends of the trait distribution, then a more diverse consumer community may impose stabilizing selection, and thus decrease the evolutionary potential of their resource. In contrast, additional consumers may impose disruptive selection if they increase the fitness impact on the mean phenotype of the resource, which would act to increase genetic variance in the trait.**

To examine these different possibilities, we conducted a field experiment that removed a consumer guild that **parasitizes**<!--directly and indirectly impacts fitness of -->an abundant insect herbivore (*Iteomyia salicisverruca*, Family Cecidomyiidae; fig. \ref{fig:Conceptual}). The larvae of this herbivore induce tooth-shaped galls when they feed on the developing leaves of willow trees [*Salix* sp., @Russo2006]. These galls protect larva from attack by generalist predators (e.g. ants, spiders), but they suffer high mortality from egg and larval parasitoids [@Barbour2016]. We manipulated food-web structure by either removing larval parasitoids (removal food web) or allowing both egg and larval parasitoids to impose selection on gall midge traits (original food web, fig. \ref{fig:Conceptual}). <!-- Omit due to Reviewer #1 comment: The relative abundance of these larval parasitoids is much lower than the egg parasitoid [@Barbour2016], and thus represents a plausible extinction scenario for this system. <!--While this manipulation simulates an extreme co-extinction event, the likelihood of such co-extinction events is increasing in the Anthropocene. -->Larval parasitoids also impose indirect effects on gall midge fitness through intraguild predation on the egg parasitoid (fig. \ref{fig:Conceptual}). We applied modern statistical methods to quantify how changes in food-web structure altered the slope and curvature of the gall midge's **fitness** landscape. Taken together, our study gives insight to how the loss of biodiversity may **alter both the trajectory and evolutionary potential**<!--constrain the evolution -->of interacting populations.

@@

We setup our food-web manipulation across 128 plants soon after galls began developing on willows in early June of 2013. These 128 plants came from **8** different plant genotypes that spanned the range of trait variation observed in this willow population [@Barbour2015]. For the original food web (eight replicates per genotype), we used flagging tape to mark 14 galled leaves per plant (~30 larvae), allowing the full suite of egg and larval parasitoids to impose selection. Marking galls with flagging tape ensured that **we compared galls with similar phenology in both treatments** when we collected galls later in the season. For the removal food web, we enclosed 14 galled leaves with 10x15cm organza bags (ULINE, Pleasant Prairie, WI, USA) to exclude three parasitoid species that attack during larval development. This treatment did not exclude the egg parasitoid *Platygaster* sp., which attacks prior to gall initiation [larva initiate gall development in Cecidomyiid midges: @Gagne1989]. It was not possible to reciprocally manipulate parasitoid attack (i.e. exclude egg parasitoid, but allow larval parasitoids), because it was not possible to identify midge oviposition sites prior to gall formation. In late August, we collected marked and bagged galls from each plant, placed them into 30 mL vials and kept them in the lab for 4 months at room temperature. We then opened galls under a dissecting scope and determined whether larvae survived to pupation (our measure of fitness) or **died due to parasitism. We further distinguished whether mortality was caused by an egg or larval parasitoid**. **Early larval death was another important source of mortality (`r round(mean(full_df$early.larval.death), 2)*100`%), but we excluded these cases from our analysis. We did this because early larval death can stunt gall growth, which would confound estimates of selection due to parasitism on chamber diameter (one of the phenotypes we measured; details given below).** 
<!--We did not include other sources of mortality in our analyses, such as early larval death, as they could influence **gall growth, which would confound estimates of selection due to parasitism on chamber diameter (one of the phenotypes we measured; details given below)**. <!--Since we were interested in selection imposed by parasitoids, we excluded other sources of mortality (`r round((mean(full_df$early.larval.death) + mean(full_df$unknown_mortality)), 2)*100`% mortality from other sources). Another reason why we excluded other sources of mortality, such as early larval death, is that they can influence the expression of the gall phenotype and confound estimates of selection.--> For the food-web treatment that excluded larval parasitoids, we further restricted our data by removing any incidental instances of parasitism by a larval parasitoid. This represented less than 3\% of the observations in this food-web treatment and allowed us to focus our inferences of selection on those imposed by the egg parasitoid. Our final dataset contains survival estimates for `r no_larvae` larvae from `r no_galls` galls and `r no_plants` plants.

@@

We collected data on three different traits that we expected to influence larval survival based on previous work in this system [@Barbour2016] and other work with gall midges [@Weis1983; @Heath2018]. First, we measured gall diameter as the size of each gall chamber to the nearest 0.01 mm at its maximum diameter (perpendicular to the direction of plant tissue growth). Previous work in this system has shown that larger galls are associated with higher survival [@Barbour2016]. Second, we measured clutch size by counting the number of chambers in each gall [@Weis1983; @Heath2018]. All larvae collected from the same multi-chambered gall were scored with the same clutch size. Third, we measured oviposition preference **for individual plants** as the density of larvae observed on a plant in an independent survey. We did this by randomly sampling five branches per tree and counting the number of individual gall chambers (number of larvae). We then converted these counts to a measure of larval density per 100 shoots by counting the number of shoots on the last branch we sampled. All larvae collected from the same plant were scored with the same oviposition preference. Measuring larval densities on plants in the field is a common method for measuring oviposition preference [@Gripenberg2010]; however, caution must be taken in inferring 'preference' as larval densities can be influenced by processes other than preference [@Singer1986]. Fortunately, **two** features of our study system suggest that larval densities may be a good proxy for oviposition preference. **First**, since our data comes from a randomized placement of plant genotypes in a common garden, there is no consistent bias in which plant genotypes females are exposed to while searching for oviposition sites. **Second**, egg predation is a minor source of mortality for galling insects in general [@Hawkins1997]; therefore, we do not expect any prior egg predation to bias our estimates of observed larval densities.

## Quantifying the **Fitness** Landscape

Inferring **fitness** landscapes assumes that trait distributions are multivariate normal [@Lande1983]. 
To approximate this assumption, we log-transformed clutch size and square-root transformed oviposition preference. **Chamber diameter already had a normal distribution so we did not transform it.** We then scaled all phenotypic traits (mean=0 and SD=1) across treatments prior to our analyses (detailed below) to ensure that our estimates of selection were comparable across traits and with other studies.

Our analysis consisted of four parts. First, we used generalized linear mixed models (GLMM) to quantify selection surfaces \textemdash linear and nonlinear relationships between absolute fitness ($W$) and standardized phenotypic traits ($i$) of individuals \textemdash in each food-web treatment. Second, we translated selection surfaces into the scale of relative fitness ($w\boldsymbol{=W/\bar{W}}$) in order to calculate standardized selection gradients. Third, we used our estimates of directional selection gradients to characterize the slope of the **fitness** landscape, which we used to measure selective constraints. Finally, we estimated the curvature of the **fitness** landscape and used a simulation to explore its effects on the adaptive potential of the gall midge population in the next generation.

**Selection surfaces**: Since larval survival was our measure of absolute fitness, we used a GLMM that assumed a binomial error distribution (and logit-link function). To approximate the selection surface, we modelled larval survival as a function of food-web treatment as well as linear ($\alpha_i$), quadratic ($\alpha_{ii}$), and linear interactions ($\alpha_{ij}$) between each trait. <!--We also allowed these selection surfaces ($\alpha$) to vary between food-web treatments.--> Note that to obtain valid estimates of linear selection surfaces, we removed nonlinear terms prior to estimating linear relationships [@Lande1983]. Other approaches have been advocated for approximating selection surfaces [@Schluter1988]; however, our approach enables us to calculate selection gradients, and thus is more appropriate for approximating the **fitness** landscape [@Arnold2003]. To account for the nonindependence of clutch size (measured at gall level) and oviposition preference (measured at plant level) as well as any independent effects of willow genotype on larval survival, we modelled gall **identity** nested within plant **identity** nested within genotype **identity** as random effects. Although statistical models with random effects are not common in analyses of natural selection, we think that modelling random effects can mitigate biased estimates of selection due to environmental covariances between traits and fitness [@Rausher1992]. Since our end goal was to characterize the relationship between mean trait values and mean fitness (**fitness** landscape), we assumed the mean value of our random effects (i.e. setting them to zero) when calculating selection surfaces. We then used parametric bootstrapping (1,000 replicates) to estimate the effect of food-web treatment on larval survival as well as selection surfaces in each food-web treatment. To determine whether trait-fitness relationships differed between food-web treatments, we calculated the difference in bootstrapped replicates between treatments.

**Selection gradients**: Selection gradients cannot be estimated directly from GLMMs of selection surfaces because the response is in terms of absolute fitness and the coefficients are on a nonlinear scale. For example, the coefficients in the **previously described binomial** model measure the change in the **log-odds** of surviving (i.e. $\text{ln}\{W(z)/[1-W(z)]\}$) associated with 1 SD change in a trait with all other traits held fixed at their mean. Therefore, we used the method developed by @Janzen1998 to translate selection surfaces from the above model into the scale of relative fitness in order to calculate directional ($\beta_i$), quadratic ($\gamma_{ii}$), and correlational ($\gamma_{ij}$) selection gradients. Briefly, this method calculates the average gradient of selection surfaces by multiplying the average of $W(z)[1-W(z)]$ by each regression coefficient (e.g. $\alpha_i$, $\alpha_{ii}$, or $\alpha_{ij}$). We then divided this average gradient by the mean fitness ($\bar W$) to put it on the scale of relative fitness, and thus interpretable as a selection gradient. We estimated selection gradients separately for each food-web treatment. We also determined whether selection gradients differed between food-web treatments by calculating the difference in bootstrapped replicates between treatments. Note that we doubled all quadratic terms prior to calculating selection gradients to put them on the same scale as estimates of directional and correlational selection [@Stinchcombe2008].

**Selective constraints**: <!--We measured evolutionary constraints by inspecting the slope and curvature of the adaptive landscape. -->The number of selective constraints is determined by the slope of the **fitness** landscape, which in our study corresponds to:

$$\textbf{Slope} = \begin{pmatrix} \beta_{\text{Diam}} \\ \beta_{\text{Clutch}} \\ \beta_{\text{Pref}} \end{pmatrix} $$
where each $\beta_i$ corresponds to the directional selection gradient acting on each trait. By comparing the number of directional selection gradients that show clear evidence of contributing to the slope (i.e. 95%  CI does not overlap zero) in our food-web treatments, we can infer the effect of food-web structure on \textbf{the number of} selective constraints. <!--This gives insight as to whether there is a specific combination of phenotypes favored by natural selection (i.e. more traits under directional selection), or whether multiple trait combinations have comparable fitness (i.e. less traits under directional selection).-->

**Genetic constraints**: The indirect effects of selection on the G-matrix ($\Delta\text{G}=\text{G}(\boldsymbol{\text{C}})\text{G}$) are governed by the curvature of the **fitness** landscape ($\text{C}=\gamma - \beta \beta^\text{T}$), which in our study corresponds to:

@@

where each $\gamma_{ii}$ (diagonal) corresponds to the quadratic selection gradient acting on a trait, and each $\gamma_{ij}$ (off-diagonal) corresponds to the correlational selection gradient acting on a particular trait combination. Note that we ommitted the upper triangle of each matrix for clarity since it is simply the reflection of the lower triangle. Subtracting these two matrices results in the curvature matrix of the **fitness** landscape:

@@

The G-matrix itself is a complex structure, but has a clear theoretical link to the evolvability of phenotypic traits [@Hansen2008]. <!--We used evolvability as our measure of genetic constraint, as it has a solid foundation in theory of quantitative genetics and adaptive landscapes [@Hansen2008]. -->Evolvability measures the ability for a trait to evolve toward a given direction of selection [@Hansen2008]. In the absence of knowledge about the direction of selection that a population will actually experience in the next generation, we can compute its average evolvability over random **directional** selection gradients [@Hansen2008; @Melo2015]. By computing the average evolvability (here, we used `r n_random_gradients` random $\beta$s) for each of `r n_Gmats` random G-matrices, we can then assess how changes in the curvature matrix alter the evolutionary potential of the associated traits. We report the distribution of these effect sizes, rather than conduct a statistical test, because the number of replicates in a simulation can be arbitrarily high, thus making any effect size statistically significant [@White2014].

@@

Once parasitized, the gall phenotype becomes the phenotype of the egg parasitoid. This phenotype may influence the egg parasitoid's survival in the face of larval parasitoids, and thus experiences selection. Our food-web manipulation allows us to measure selection imposed by larval parasitoids on the phenotype of egg parasitoids. Using the same models as described above, we substituted egg parasitism as our response variable to quantify selection surfaces and selection gradients acting on the egg parasitoid. Note that we cannot test the effect of food-web structure on the egg parasitoid's **fitness** landscape ---we can only estimate selection imposed by larval parasitoids. This comparison is still useful though in determining the extent to which the loss of consumers may have indirect evolutionary effects by altering selection on multiple interacting populations.

All analyses and visualizations were conducted in R [@R2018]. Unless otherwise noted, we report mean estimates of selection surfaces and selection gradients with 95% confidence intervals in brackets. Note that for visualizing the **fitness** landscape we restrict trait axes to $\pm 1$ SD of the mean trait value. This emphasizes the fact that we can only reliably estimate the shape of the **fitness** landscape near the mean phenotype of the population [@Arnold2001]. We also plot mean larval survival on a natural log scale to accurately reflect the mathematical definition of the **fitness** landscape [@Arnold2003]. All data and code to reproduce the reported results are publicly available on GitHub (https://github.com/mabarbour/complexity_selection) and have been archived on Zenodo (https://zenodo.org/badge/latestdoi/108833263).
